---
title: Habitat Selection by Cottonmouths (Agkistrodon piscivorus) in Middle Tennessee.-
  Preliminary Anlysis
author: "Jack Kauphusman"
date: "10/23/2019"
output:
  html_notebook:
    df_print: paged
    highlight: breezedark
    number_sections: yes
    rows.print: 10
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document: default
editor_options:
  chunk_output_type: inline
---
<style type="text/css">

h1.title {
  font-size: 40px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
  font-size: 20px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
</style>
---


```{r Library, include=FALSE}
library("dplyr")
library("ggplot2")
library("factoextra")
library("FactoMineR")
library("tidyr")
library("knitr")
library("gridExtra")
library(lattice)
library(permute)
library(vegan)
library(RVAideMemoire)
library(MASS)
library(caret)
library(tmap)
library(tmaptools)
library("ggsn")
library("leaflet")
library("mapdata")
library("maptools")
library("rgdal")
library("smoothr")
library("sf")
library("sp")
library(ggmap)
library("tidyverse")
library("randomForest")
library("randomForestExplainer")
library("rfUtilities")
library(faraway)
library(ROCR)
library(formattable)
library(kableExtra)
```

# **Methods**

## Study Area 

Data was collected along the Cumberland River Bicentennial Trail (CRBT), Cheatham County, Tennessee, USA (36˚17’ N, 87˚05’W). A boundary of an 115ha area was constructed using ArcMap (ESRI Products) to be used as the focal study area and the habitats within to be assessed. The trail was included within the boundary and divided the mixed lowland forest and slough wetlands from the upland limestone bluffs and rocky outcrops. Cottonmouths used the upland forest and limestone bluffs as overwintering sites and occupied the lowland forest slough during the active season. The landscape features that bordered the study area were the limestone bluffs and backwater streams along the northern portions, the Cumberland River along the west and south, and crop fields and urban development along the southern portions. These landscape features appeared to isolate this population of cottonmouths from dispersal. Unpublished telemetry data (Smith 2009) noted that snakes did not leave the study area, however, an adult male was observed attempting to cross a road toward the Cumberland River but was unfortunately killed by a passing vehicle.  

The habitat types found within the study area are the buttonbush slough in the central portion (xx ha), characterized by relatively open canopy, deeper water (50-150 cm), and mats of floating vegetation (Hydrocotyle ranunculoides & Ludwigia palustris) scattered throughout the slough. An open water habitat (> 100cm) is located toward the western portion of the study area. The seasonally flooded lowland forest (xx ha), dominated by red maples (Acer rubrum), has the denser canopy closure for the wetland, and is located between the slough and upland forest. Toward the southern portion between the slough and crop fields is a transitional edge area (xx ha) dominated by dense grass vegetation and occasionally floods (<10cm) in some portions throughout the year.  The upland forest habitat (xx ha) was dominated by oak trees (Quercus sp.) and had dense canopy closure and higher percentage of bare soil. Three main streams flow into the slough from the upland forest and provide a permanent source of water to the slough, along with periodic flooding from the Cumberland River, and groundwater seepage. A constructed levee divides the slough from adjacent crop fields, and during the fall waterfowl hunt, a culvert is opened from the levee to reduce the flooded areas.  

```{r Field Map, echo=TRUE, message=FALSE, warning=FALSE}
#Outmap
state <- map_data("state")
county <- map_data("county")
study_site <- data.frame("x" = -87.092447, "y" = 36.288684)
tn <- county %>% 
  filter(region=="tennessee")

Cheatham <- county %>% 
  filter(region=="tennessee") %>% 
  filter(subregion=="cheatham")

a1<-ggplot() + geom_polygon(data = state, aes(x=long, y = lat, group = group),
                        fill = "white", color="black") + 
           geom_polygon(data = tn, aes(x=long, y = lat, group = group),
                        fill = "gray", color="black") + 
           geom_polygon(data = Cheatham, aes(x=long, y = lat, group = group),
                        fill = "yellow", color="black") + 
           geom_point(data = study_site, aes(x=x,y=y), color="black") +
  coord_fixed(xlim = c(-91, -81),  ylim = c(34, 37), ratio = 1.2) + 
  xlab("Longitude") + ylab("Latitude") + ggtitle("Cheatham Co., TN")
```

```{r Files, include=FALSE}
#Using ggspatial
library(ggspatial)
library(rosm)

#Shapefiles
generalhab<-st_read("data/Habitat/GENERALHABITAT.shp")
Boundary<-st_read("data/Habitat/GENRLHABI.shp")
cottonmouths<-st_read("data/Habitat/Cottonmouths.shp")
```


```{r Study Site Static, echo=TRUE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
#Map
b<-ggplot()+
  annotation_map_tile(type = "thunderforestlandscape")+
  annotation_scale(location = "bl")+
  geom_sf(data = Boundary,size = 1, col = "red", fill = NA, show.legend = "line")+
  geom_sf(data = cottonmouths, aes(fill = "Cottonmouth Location"),col = "purple",size = 1.5, show.legend = "point")+
  labs(fill = 'Legend', title = "Study Site")
grid.arrange(a1,b, ncol=1)
```
**Map 1:** Above, location of the study site along a swamp adjacent to The Cumberland River Bicentennial Trail, Cheatham County, TN, USA. Below,is the zoomed in view of the study area. The boundary line, where collection took place, is colored in red, and the cottonmouth locations were symbolized as green points. 

```{r Study Site, echo=TRUE, message=FALSE, warning=FALSE}
a<-tmap_mode("view")+
tm_shape(Boundary)+
  tm_borders(col= "red", lwd = 1)+
  tm_shape(cottonmouths)+
  tm_dots(col = "purple", size = 0.05, id = "NAME", text = "Cottonmouth location")+
  tm_scale_bar()+
  tm_layout("Study Site", legend.show = T)+
  tm_basemap(leaflet::providers$Esri.WorldImagery)
a
```
**Map 2:** The interactive map of the study area, made using the tmap package.


## Field Collection 

Visual encounter surveys occurred during the active seasons (May – October) of 2012, 2013, and 2019. Snakes were captured using snake tongs and retained in plastic restraining tubes (Midwest Tongs).  For individual identification, each snake was implanted subcutaneously with a Passive Integrated Transponder (PIT) tag (Biomark) on the posterior to the ventral side of the midbody. Tags and implant syringes were soaked in Benzalkonium Chloride (Benz-all) for 2 – 3 minutes prior to tag implantation, and an antiseptic liquid bandage (Vetclose) was applied to the insertion area to facilitate healing. Snout-vent length (SVL) and tail lengths (TL) were measured, and snakes were sexed by either everting hemipenes or by probing (Blanchard, F and Finister, E 1933) Individuals were considered juveniles if their SVL was between 500 - 300mm (Burkett, Ray 1966, Eskew et al. 2009), and young of the year (YOY) were considered to be individuals less than 300mm SVL (Koons et al. 2009). Body mass was measured with either a spring scale (Pesola) in the field, or by digital balance (Ohaus) in the lab. 

Habitat selection was quantified by recording a suite of biotic and abiotic environmental variables (Table 1) from the location an individual was located prior to capture. A 1-m2 quadrat was used as the scale at which habitat variables were measured. To assess the microhabitats available to the population within the study site to those habitat variables being selected by cottonmouths, a combination of random and fixed sites (habitat sites) were measured. A randomly chosen comparison site was paired with each snake location, and sites were selected by using the random points tool in ArcMap within the boundary area. Fixed sites were placed 200-m in relation to one another within the boundary area, using the Fishnet tool in ArcMap, to assure all available habitats were quantified. 

## Multivariate Analysis 

To test that cottonmouth habitat selected was different then general habitat available a Permutational Multivariate Analysis of Variance (PERMANOVA) on Euclidean distance was performed (Anderson 2017). PERMANOVA tests for discrimination of groups within multivariate space, and a significant p-value (<0.05) indicates that groups differ due to their distances between centroid locations, their relative dispersion, or both (Anderson 2017). The advantages of PERMANOVA are that it does not assume normality and homogeneity of variance (Delgado-Baquerizo et al. 2015). Permutational analysis of multivariate dispersion (PERMDISP) tests for homogeneity of multivariate dispersion among groups, and is used to distinguish if dispersion is causing a significant PERMANOVA(Assis et al. 2013, Anderson 2017, Brückner and Heethoff 2017). A significant p-value from PERMDISP indicates that multivariate dispersion is heterogeneous between groups (Anderson 2017). PERMANOVA/PERMDISP were computed in R using the vegan package (Oksanen et al. 2019). To visualize the differences between cottonmouth habitat and general habitat, and aid in interpretation of the PERMANOVA/PERMDISP results, a Principal Component Analysis (PCA) on correlations was used. PCA helps to reduce dimensionality of a large dataset by extracting the most contributing variables that explain the variation within each dimension, and rank components based on explained variance (Peres-Neto et al. 2003). Prior to PCA, habitat variables were scaled (Kassambara 2017), and the first two components were retained for better visual representation. PCA was performed using the R packages FactoMiner and FactoExtra (Kassambara 2017).  

To test for habitat selection differences between cottonmouth groups (Males, Females, Juveniles and YOY) PERMANOVA/PERMDISP and PCA were used. A pairwise comparison of PERMANOVA was used to identify differences between each group (Herve 2019). To evaluate the presence of an ontogenetic difference in overall habitat use among cottonmouths, linear regression was used to compare snake SVL and principal component one, similar to Eskew et al. 2009. 

## Random Forest  

Random forest (RF) classification approach (Breiman 2001) was used to identify the important habitat variables that discriminate between cottonmouth selected habitat and general habitat available. RF is a nonparametric classification method that builds an ensemble of thousands of classification trees from predictive variables and optimizes the accuracy of each tree. For each tree, a bootstrap sample of approximately 64% (i.e., bag sample) is used as a training set, and the remaining data left out is used as a validation set (i.e., out-of-bag [OOB]) to the finished tree as a cross-validation method (Breiman 2001, Severson et al. 2017). RF uses all trees to predict and estimate the overall OOB error rate. RF classifiers are resistant to collinearity of covariates, do not assume independence of samples, model nonlinear relationships, and are known to be more accurate models with ecological data compared to general statistical (ex: logistical regression) (Breiman 2001, Cutler et al. 2007, Severson et al. 2017).  However, multi-collinearity can reduce interpretations of results (Murphy et al. 2010). Before model construction, multi-collinearity in the predictors were assessed using qr matrix decomposition (threshold = 0.05) using the rfUtilities package (Evans and Murphy 2019). Using the randomForest function, from the randomForest package (Breiman et al. 2018), 1,000 classification trees were applied to each model. To construct an optimal model classifying cottonmouth and general habitat from the collected habitat variables, 30 iterations of a 10-fold cross-validation training set were ran, and a mean RF model was calculated. The cross-validation test set was then used to calculate a mean error rate and identify a minimal error rate model (Severson et al. 2017). The minimal error rate model was used to calculate relative importance and create partial dependence plots of the top 4 most contributing predictive habitat variables used to discriminate groups in the final model. 

# **Results**

```{r General Stats, include=FALSE}
s<-read.csv("data/snake.SVL.csv")
s<-na.omit(s)
s<-s[,2:28]
#How Many snakes (in groups) and general habitat variables measured

Female<-tabulate(s$Type)[1]
Juvenile<-tabulate(s$Type)[2]
Male<-tabulate(s$Type)[3]
YOY<-tabulate(s$Type)[4]
Snake<- Female+Juvenile+Male+YOY

sg<-sg<-read.csv("data/HabitatSel.csv")
General<-tabulate(sg$Type)[1]

#SVL 

SVL<-group_by(s, Type)%>%
  summarise(
    mean=mean(SVL, na.rm = T)
  )
SVl.SD<- group_by(s, Type)%>%
  summarise(
    sd=sd(SVL, na.rm = T)
  )
```

From May - October  of 2012, 2013, and 2019, we recorded habitat variables for `r Snake` cottonmouths and `r General` general habitat comparison sites. Of the cottonmouths that were captured, `r Male` were adult males, `r Female` adult females, `r Juvenile` juveniles, and `r YOY` YOY. Adult males averaged in **(744 +/-132)**mm in SVL, females **(620 +/- 60)**mm, juveniles **(408 +/- 55)**mm, and YOY **(251 +/- 24)**.  

## Cottonmouth Habitat Compared To General Habitat.

We found that cottonmouth habitat selection significantly differed from general habitat available to the cottonmouths within the study area (PERMANOVA; P< 0.001; Pseudo-F = 19.9; d.f. = 1; Table 1). The PERMDISP test found that there was equal dispersion among groups and no significant difference between group centriods(PERMDISP; P=0.17; F = 1.90; Table 1), thus differences between the groups was due to heterogenous dispersion within groups in multivariate space. The Principal component analysis (PCA-1) comparing cottonmouth habitat compared to general habitat explained 30.3% of the variation within dimensions one and two. The most contributing variables within each dimension in PCA-1, were Percent Land (19%), Percent Water (19%), Distance to Water (13%), and Water Depth (13%) within dimension 1, and Ambient Temp (24%), Surface Temp (26%), and Substrate Temp (20%) within dimension 2 (Fig 1a). The results of the PERMANOVA/PERMDISP align with the observations from PCA1 in that the heterogenous dispersion within each group show that cottonmouth habitat was within the general habitat ellipse (Fig 1a). 

## Habitat Analysis on Cottonmouth Groups

The results from the multivariate analysis investigating habitat selection differences among cottonmouth groups found that there were no significant differences between groups (PERMANOVA; P >0.05; Pseudo-F =1.21; Table 2). There was homogenous distribution of variance among groups present in multivariate space and no significant difference between centroids (PERMDISP; P> 0.05, F = 2.15; Table 2). A parwise comparision on the PERMANOVA using Hotelling T-square test identified that there was no significant difference between groups habitat selection. Although, the pairwise comparison of adult males and YOY were close to being significantly different (Pairwise PERMANOVA; P = 0.08). The Principal component analysis (PCA-2) comparing habitat selection among the cottonmouth groups explained  31.1% of the variation within dimensions one and two. Habitat variables with the highest contributions were shared with PCA1. Percent Land (17%) and Percent Water (17%) had the highest contribution to dimension 1, and Ambient Temp (23%), Surface Temp (26%), and Substrate Temp (20%) in dimension 2.

Linear regression comparing dimension one habitat variables of PCA-2 and body size (SVL) found a significant negative relationship (F = 3.29;P <0.01;d.f. = 1; Fig 2). The linear relationship indicated that larger snakes prefered wetter habitat sites along the dimension one, which was a gradient from percent land variables to percent water variables. Caution must be used in interpreting the results as this relationship had a low explainatory power in dimension one (20%) and R2 (0.02) value.

## Random Forest

Before the Random Forest models were ran, we used a VIF test on the twenty-three habitat variables, and found that percent land was multicollinear, so it was excluded from the models.  The 30-iteration random forest models, used to identify cottonmouth habitat selection from habitat available within the study site, had a mean accuracy of 81.9% from cross-validation and an 83.1% mean accuracy from the OOB (Table 3).  In addition, the mean of the random forest model’s accuracy metrics had a 63.8% kappa (a measure of agreement between predicted presences and absences with actual presences and absences corrected for agreement that might be due to chance alone), 92.9% sensitivity (the percentage of presences correctly classified), 71.0% specificity (the percentage of absences correctly classified), and 89.8% for the area under the receiver operating characteristic curve (AUC) (Table 3) (Cutler et al. 2007). Maximum cross-validation accuracy was 82.1% and 89.3% OOB accuracy for the RF models, and the minimum cross-validation accuracy was 78.6% and 82.4% OOB (Table 3). Random forest models error rates stabilized at approximately 3,000 trees, indicating that 5,000 trees were adequate. 

Our maximum accuracy RF model identified that distance to water (cm) and water depth (cm) were the most important predictive variables to identify the occurrence of a cottonmouth (Fig 2). The probability of occurrence of a cottonmouth decreased as you increased in distance to water within the study area (Fig 3A). Response to water depth sharply increased from 0 – 15 cm and then decreased in occurrence as depth increased past 25 cm (Fig 3B). Response to water cover increased in occurrence and then gradually leveled past 10% cover (Fig 3C). Response to bare soil cover increased up to 30% and then slightly declined in cottonmouth occurrence (Fig 4D).  



## Result Tables & Figures

```{r Snake Vs. Random PCA, include=FALSE}
s<-read.csv("data/HabitatSel.csv")
s<-na.omit(s)
s<-s[,c(1:24)]
```

```{r PERMANOVA & PERMDISP Tests}
#PERMANOVA
Perm1<-adonis(s[,2:24]~ Type, data = s, method='eu')
Perm1<-cbind(Perm1$aov.tab[1,4], Perm1$aov.tab[1,6])

#PERMDISP
dist<-dist(scale(s[,2:24]))
bd<-anova(betadisper(dist,s$Type))
bd1<-cbind(bd$`F value`, bd$`Pr(>F)`)
bd1<-data.frame(bd1)
bd1<-bd1[-2,]

PermDisp<- cbind(Perm1, bd1)
names(PermDisp)<-c("Pseudo-F", "P", "F", "P ")
row.names(PermDisp)<-c("Proportions (Euclidean)")

PERM<- kable(PermDisp, digits = 4, caption = "Table 1: Results of the multivariate statistal tests PERMANOVA/PERMDISP based on the Euclidean distance comparing cottonmouth habitat selected and general habitat available to the cottonmouth within the study area.")%>%
  kable_styling(c("striped", "bordered"))%>%
  add_header_above(c(" " = 1, "PERMANOVA" = 2, "PERMDISP" = 2))%>%
  footnote(general = "N(permutations) = 999; df = 1")

PERM
```
```{r Dataset PCA}
#Uploading Data
s<-read.csv("Data/CottonmouthTotal-11-18.csv")
#s<-s[c(1:150),c(3:26)]
s<-na.omit(s)
```


```{r PCA, fig.width=15, fig.height=10}
#PCA
s1<-s

pc<-princomp(s1[,3:25], cor = T, scores = T)
summary(pc)
#Component
#Setting- Up ANOVA
comp<-pc$scores[,1:3]
s1<-cbind(s1, comp)

#Snake Averages
s2<-s1[c(1:149),c(3:32)]
pcomp<-group_by(s2, Group)%>%
  summarise(Comp.1 = mean(Comp.1),
            Comp.2 = mean(Comp.2),
            Comp.3 = mean(Comp.3))

#Component Percentage
summary(pc)


#Plot
p<-ggplot(data = s1, aes(Comp.1, Comp.2))+
  geom_point(alpha = 0.5, size = 1.5)+
  geom_point(data = pcomp, aes(pcomp$Comp.1, pcomp$Comp.2, shape= Group), size = 5)+
  #geom_label_repel(data = pcomp, aes(label = Group), size= 5, guide_legend= F)+
  theme_bw(base_size = 24)+
  theme(panel.grid = element_blank(), panel.border = element_rect(fill= "transparent"))+
  geom_hline(yintercept = 0, lty = 2) +
  geom_vline(xintercept = 0, lty = 2)+
  stat_ellipse(type = "norm", aes(color = Type), 
               show.legend = FALSE,size = 1, level = 0.95, linetype = 2)+
  scale_color_manual(values = c("black","#636363"))+
  guides(shape= guide_legend("Cottonmouth"))+
  geom_text(aes(-2.5,2.2), label = paste("Cottonmouth"), col = "#636363", size = 9)+
  geom_text(aes(-2.7,4), label = paste("General"), col = "black", size = 9)+
  xlab(expression(atop("flooded sites" %<->% "non-flooded sites", paste("PC-1 (19.6%)" ))))+
  ylab(expression(atop ("PC-2 (12.8%)", paste("lower temp sites" %<->% "higher temp sites"))))+
  theme(legend.position=c(.9, 0.88), legend.background = element_rect(fill="White",
      size=0.5, linetype="solid", colour ="black"))
p
```

**Figure 1:** Principal component analysis of the environmental variables measured for the *A. piscivorus* (cottonmouth) and general available habitat selected comparison sites. Ellipses represent 95% condifence intervals for cottonmouths (grey) and general habitat (black). Mean principal component scores of the intra-specific cottonmouth groups (adult males, adult females, juveniles,  & young of the year [YOY]) plotted. Positive values for PC-1 indicate non-flooded areas with a greater distance from flooded areas, and positive PC-2 values indicate areas with higher thermal sites.

```{r ANOVA, echo=TRUE, message=FALSE, warning=FALSE}
pcomp<-pcomp[,2:4]
names(pcomp)<- c( "PC-1","PC-2", "PC-3")
row.names(pcomp)<-c("Female", "Juvenile", "Male", "YOY")
pcomp[,1:3]<-round(pcomp[,1:3], digits = 3)


C1<-aov(Comp.1~Group, s2)
C2<-aov(Comp.2~Group, s2)
C3<-aov(Comp.3~Group, s2)


F.value<-c(anova(C1)$"F value"[1], anova(C2)$"F value"[1], anova(C3)$"F value"[1])
p.value<-c(anova(C1)$"Pr(>F)"[1],anova(C2)$"Pr(>F)"[1],anova(C3)$"Pr(>F)"[1])

stats<-data.frame(t(cbind(F.value, p.value)))
names(stats)<-c( "PC-1","PC-2", "PC-3")
pcomp<-rbind(pcomp,stats)

row.names(pcomp)<-c("Female", "Juvenile", "Male", "YOY", "F", "P-value")

ANOVA<-kable(pcomp, row.names = TRUE, col.names=c( "PC-1","PC-2", "PC-3"),digits = 3, align = "l", caption = "Table 2: Mean component scores of PCA on intra-specific habitat selection within 
      cottonmouths with ANOVAs used to compare groups.")%>%
  kable_styling(c("striped", "bordered"))%>%
  pack_rows("ANOVA", 5,6,latex_gap_space = "2em")

ANOVA
```

```{r Tukey HDS Test}
TukeyHSD(C1)

TukeyHSD(C2)

TukeyHSD(C3)
```


```{r Envi & Loading Scores, echo=TRUE, message=FALSE, warning=FALSE}
Description<- c("Temperature within 2 cm into the substrate or water","Temperature at surface level","Temperature at breast height", "Length of cover object closest to snake point","Width of cover object closest to snake point", "Percentage of canopy within quadrat", "Coverage within quadrat","Coverage within quadrat","Coverage within quadrat",
"Coverage within quadrat", "Diameter of largest living woody stem", "Height of largest living woody stem", "Height of cover object closest to snake point", "Distance to nearest cover object", "Diameter of largest log within quadrat", "Distance to nearest water source", "Depth measured closest to snake point", "Coverage within quadrat","Coverage within quadrat","Coverage within quadrat","Coverage within quadrat", "Coverage within quadrat","Coverage within quadrat", "", "")

loadings<-pc$loadings[,1:3]
sdev<-t(data.frame(pc$sdev))
eigen<-sdev[,1:3]
Variance<-c(19.61, 12.83, 9.71)
loadings<-rbind(loadings, eigen, Variance)
loadings<-round(loadings, digits = 2)
names(loadings)<-c("PC-1", "PC-2", "PC-3")
loadings<-cbind(Description, loadings)

row.names(loadings)<-c("Substrate Temp (◦C)","Surface Temp (◦C)", "Ambient Temp(◦C)","Cover Length(cm)", "Cover Width (cm)", "Canopy Closure (%)", "Rock Cover (%)", "Leaf Cover (%)", "Vegetative Cover (%)", "Fallen Log Cover (%)", "Woody Stem Diameter (cm)","Woody Stem Height (cm)", "Cover Height (cm)", "Distance To Cover (cm)","Log Diameter (cm)", "Distance To Water (cm)", "Water Depth (cm)", "Percent Land (%)", "Percent Water (%)", "Percent Woody Debris (%)", "Percent Debris (%)", "Percent Floating Vegetation (%)", "Percent Bare Soil (%)", "Eigenvalue", "Explained Variance")

load<-kable(loadings, row.names = TRUE,digits = 2, col.names = c("Description", "PC-1", "PC-2", "PC-3"), align = "l", booktabs = T, caption = "Table 1: Factor loadings from principal component analysis of environmental variables measured for cottonmouths and general habitat comparitsion sites")%>%
  kable_styling(c("striped", "bordered"))%>%
  pack_rows("",24,25,latex_gap_space = "2em")

load
```


```{r Regression Line, fig.width = 10,echo=TRUE}
#Linear Regression
mod<-lm(SVL~Comp.1,s2)
summary (mod)

mod2<-lm(SVL~Comp.2, s2)
summary(mod2)

PC1<-ggplot(s2, aes(x= SVL, y= Comp.1))+
  geom_point(size = 2, aes(shape = Group))+
  xlim(209, 1061)+
  geom_smooth(method = "lm", color= "black")+
  geom_hline(yintercept = 0, lty = 2) +
  geom_vline(xintercept = 600, lty = 2)+
  theme_bw(base_size = 24)+
  geom_text(aes(1000, 6, label = paste("R-squared = 0.03")), color="black")+
  geom_text(aes(1000, 5.4, label = paste("F= 5.04; P= 0.03")), color = "black")+
  xlab("Snout-vent length (mm)")+
  ylab(expression(atop("flooded sites" %<->% "non-flooded sites", paste("PC-1 (19.6%)" ))))

PC1
```
**Figure 2:** Linear regression of Principal Component scores of PC-1 comparison of habitat selection among the intra-specific cottonmouth groups(Males, Females, Juveniles, and YOY) and their snout-vent lengths. 

```{r randomForest, include=FALSE}
#Loading data
s<-read.csv("data/HabitatSel.csv")
s<-na.omit(s)
s<-s[,c(1:24)]
```

```{r Checking Multicolinearrity, include=FALSE}
corrrel<-cor(s[,2:24])
findCorrelation(corrrel, cutoff = 0.7)

#OR with VIF
model<-lm(Type~., data = s)
vif(model)

#From the VIF the X.Land and X.Water have high degree of multicollinearity so I shall remove X.Land
s<-s[,c(-19,-2,-4)] #Removed X.Land, and Substrate Temp, and Surface Temp
model<-lm(Type~., data = s)
vif(model)
cor(s[,2:21])
findCorrelation((cor(s[,2:21])), cutoff= 0.75)
```

```{r Checking MIR, eval=FALSE, include=FALSE}
#Model Improvement Metric
RF<-randomForest(Type~.,s, ntree= 5000)
RF
varImp(RF, scale = F)
```


```{r 10-Fold Cross Validation, echo=TRUE}
set.seed(123)
ind<-sample(2,nrow(s),replace = T, prob = c(0.9,0.1))
train<-s[ind==1,]
test<-s[ind==2,]
```

```{r 30 iterations, echo=TRUE}
a1<- randomForest(Type~., data = train, ntree = 5000)
a2<- randomForest(Type~., data = train, ntree = 5000)
a3<- randomForest(Type~., data = train, ntree = 5000)
a4<- randomForest(Type~., data = train, ntree = 5000)
a5<- randomForest(Type~., data = train, ntree = 5000)
a6<- randomForest(Type~., data = train, ntree = 5000)
a7<- randomForest(Type~., data = train, ntree = 5000)
a8<- randomForest(Type~., data = train, ntree = 5000)
a9<- randomForest(Type~., data = train, ntree = 5000)
a10<- randomForest(Type~., data = train, ntree = 5000)
b1<- randomForest(Type~., data = train, ntree = 5000)
b2<- randomForest(Type~., data = train, ntree = 5000)
b3<- randomForest(Type~., data = train, ntree = 5000)
b4<- randomForest(Type~., data = train, ntree = 5000)
b5<- randomForest(Type~., data = train, ntree = 5000)
b6<- randomForest(Type~., data = train, ntree = 5000)
b7<- randomForest(Type~., data = train, ntree = 5000)
b8<- randomForest(Type~., data = train, ntree = 5000)
b9<- randomForest(Type~., data = train, ntree = 5000)
b10<- randomForest(Type~., data = train, ntree = 5000)
c1<- randomForest(Type~., data = train, ntree = 5000)
c2<- randomForest(Type~., data = train, ntree = 5000)
c3<- randomForest(Type~., data = train, ntree = 5000)
c4<- randomForest(Type~., data = train, ntree = 5000)
c5<- randomForest(Type~., data = train, ntree = 5000)
c6<- randomForest(Type~., data = train, ntree = 5000)
c7<- randomForest(Type~., data = train, ntree = 5000)
c8<- randomForest(Type~., data = train, ntree = 5000)
c9<- randomForest(Type~., data = train, ntree = 5000)
c10<- randomForest(Type~., data = train, ntree = 5000)
```

```{r Error Rate Plot, eval=FALSE, include=FALSE}
#Error rate of Random Forest
plot(a1)
```


```{r Out-of-Bag}
OOB1<-c(a1$err.rate[1000,1],a2$err.rate[1000,1],a3$err.rate[1000,1],a4$err.rate[1000,1],a5$err.rate[1000,1],a6$err.rate[1000,1],
       a7$err.rate[1000,1],a8$err.rate[1000,1],a9$err.rate[1000,1],a10$err.rate[1000,1],b1$err.rate[1000,1],b2$err.rate[1000,1],
       b3$err.rate[1000,1],b4$err.rate[1000,1],b5$err.rate[1000,1],b6$err.rate[1000,1],b7$err.rate[1000,1],b8$err.rate[1000,1],
       b9$err.rate[1000,1],b10$err.rate[1000,1],c1$err.rate[1000,1],c2$err.rate[1000,1],c3$err.rate[1000,1],c4$err.rate[1000,1],
       c5$err.rate[1000,1],c6$err.rate[1000,1],c7$err.rate[1000,1],c8$err.rate[1000,1],c9$err.rate[1000,1],c10$err.rate[1000,1])

OOB<-c(1-min(OOB1),1-mean(OOB1), 1-max(OOB1))
```



```{r}
files<-list(a1,a2,a3,a4,a5,a6,a7,a8,a9,a10,b1,b2,b3,b4,b5,b6,b7,b8,b9,b10,c1,c2,c3,c4,c5,c6,c7,c8,c9,c10)

cv<-(function(x){
p<-predict(x, test)
cm<-confusionMatrix(p, positive = "SNAKE",test$Type)

CVTABLE<-t(data.frame(cm$overall))
CV2TABLE<-t(data.frame(cm$byClass))

cv_table<-cbind(CVTABLE,CV2TABLE)
cv_table<-cv_table[,c(1,2,8,9)]
cv_table
})

cv_table<-t(data.frame(lapply(files, cv)))

cv_max<-c(max(cv_table[,1]), max(cv_table[,2]), max(cv_table[,3]), max(cv_table[,4]))
cv_mean<-c(mean(cv_table[,1]), mean(cv_table[,2]), mean(cv_table[,3]), mean(cv_table[,4]))
cv_min<-c(min(cv_table[,1]), min(cv_table[,2]), min(cv_table[,3]), min(cv_table[,4]))

cv_table<-data.frame(rbind(cv_max,cv_mean, cv_min))

RandomForest_Model<-c("Maximum", "Mean", "Minimum")
```

```{r Calculating AUC}

AUC<- function(z){
pred<-predict(z, test, type = "prob")
pred<-pred[,2]
pred<-prediction(pred, test$Type)

#AUC (Area Under The Curve)
auc<-performance(pred, "auc")
auc<-unlist(slot(auc, "y.values"))
  
}

AUC<-t(data.frame(lapply(files, AUC)))

AUC<-c(max(AUC), mean(AUC), min(AUC))
```


```{r Table}
models<-cbind(RandomForest_Model, cv_table, OOB, AUC)

kable(models, row.names = FALSE, col.names=c("Random Forest Model", "Accuracy", "Kappa", "Sensitivity", "Specificity", "OOB", "AUC"),digits = 3, caption = "Table 3: Performance metrics for the 30-iteration random forest model on cottonmouth habitat selection compared to general habitat available within the study area.")%>%
  kable_styling(c("striped", "bordered"))%>%
  footnote(general = "N(trees) = 5000;",
           "OOB = Out of Bag Accuracy",
           "AUC = Area Under the Curve")
```

```{r Variable Importance, include=FALSE}
#Maximal Model was a1 in the iterations, so I will now use that to create the partial - least square biplots of the top 4 most contributing variables,

#varImp(c10) Yields Similar to GINI but is relative importance

Gini<-importance(a3, class = "SNAKE")

Gini<-data.frame(Gini)
Habitat_Variables <- c("Surface Temp", "Cover Length", "Cover Width", "Canopy Closure", "Rock Cover", "Leaf Cover", "Vegetative Cover", "Fallen Log Cover", "Woody Stem Diameter","Woody Stem Height", "Cover Height", "Distance To Cover","Log Diameter", "Distance To Water", "Water Depth", "Percent Water", "Percent Woody Debris", "Percent Debris", "Percent Floating Vegetation", "Percent Bare Soil")
D<-cbind(Habitat_Variables,Gini)

G<-ggplot(D, aes( x = reorder(Habitat_Variables, MeanDecreaseGini), y=MeanDecreaseGini))+
  geom_col(width = .4)+
  ylim(0,25)+
  coord_flip()+
  xlab("Environmental Variables")+
  ylab("Relative Importance")+
  theme_bw()


#Partial Dependence Plot
w<-partialPlot(c10, train, Dist..To.Water, "SNAKE", plot = FALSE)
x<-partialPlot(c10, train, Depth, "SNAKE", plot = FALSE)
y<-partialPlot(c10, train, X.Water, "SNAKE", plot = FALSE)
z<-partialPlot(c10, train, X.Bare.soil, "SNAKE", plot =FALSE)


#Distance to Water
w<-cbind(w$x,w$y)
w<-data.frame(w)
names(w)<-c("x","y")

w1<-ggplot(data = w, aes(x,y))+
  geom_line(size = 1)+
  xlab("Distance to Water (cm)")+
  ylab("Relative probability")+
  geom_text(aes(1000,0.40), label = paste("A"), col = "black")+
  ylim(-0.8, 0.6)+
  xlim(0,1000)+
  theme_bw()

#Depth

x<-cbind(x$x,x$y)
x<-data.frame(x)
names(x)<-c("x","y")

x1<-ggplot(data = x, aes(x,y))+
  geom_line(size = 1)+
  xlab("Water Depth (cm)")+
  ylab("Relative probability")+
  geom_text(aes(80,0.23), label = paste("B"), col = "black")+
  theme_bw()
#X. Bares Soil
y<-cbind(y$x,y$y)
y<-data.frame(y)
names(y)<-c("x","y")

y1<-ggplot(data = y, aes(x,y))+
  geom_line(size = 1)+
  xlab("Water Cover (%)")+
  ylab("Relative probability")+
  geom_text(aes(100,0.47), label = paste("C"), col = "black")+
  theme_bw()

z<-cbind(z$x,z$y)
z<-data.frame(z)
names(z)<-c("x","y")

z1<-ggplot(data = z, aes(x,y))+
  geom_line(size = 1)+
  xlab("Bare Soil Cover (%)")+
  ylab("Relative probability")+
  geom_text(aes(100, 0.17), label = paste("D"), col = "black")+
  theme_bw()

```

```{r Variable importance Plot}
G
```
**Figure 3:**The Relative Importance (based on the Mean decrease in Gini) on the environmental variables from the Maximum Random Forest model on Cottonmouth (Agkistrodon piscivorus) habitat selection in Middle TN, USA. Cover variables were measured in percentage of occurance within a 1-m^2 quadrat, and all other metrics were measured in cm.

```{r Partial Dependance Plots}
grid.arrange(w1,x1,y1,z1, ncol= 2)
```
**Figure 4:** Partial dependance plots of the four most constributing variables (From mean decrease in gini) with relative probability of cottonmouth habitat selection of a site within Middle Tennessee, USA. Water depth (A), distance to water (B), percent bare soil (C), and surface temperature (D) were measured within a 1-m^2 quadrat from were a cottonmouth was located during visual encounter.

# Discussion

# References:

**FactoMiner & FactoExtra** 

Kassambara, A. 2017. Practical Guide to Principal Component Methods in R. STHDA.

**randomForest & randomForestExplainer** 

Breiman, L. & Cutler, A. 2018. Breiman and Cutler's Random Forests for Classification and Regression. CRAN.

**vegan**

Oksanen, J., Guillaume Blanchet, F., Kindt, R., Legendre, P., McGlinn, D., Minchin, P., O'Hara, R.B., Simpson, G., Solymos, P., Henry, M., Stevens, H., Szoecs, E., Wagner, H. 2019. Community Ecology Package. CRAN.
